### 2018-03-16 paper/3-abstractions.tex
Redo the Shake code, flip some arguments

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
import Control.Monad.Writer
\end{minted}
\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{haskell}
track :: Task Monad k v -> Store i k v -> k -> Maybe (v, [k])
track task store = runWriter . task (\k -> writer (getValue store k, [k]))
\end{minted}
\vspace{1mm}

\noindent
The implementation uses the standard Haskell \hs{Writer} monad, which allows us
to record addition information -- a list of keys of type \hs{[@@k]} -- when
computing a task. We instantiate \hs{f} to \hs{Writer}~\hs{[@@k]}~\hs{v}, and feed
\hs{fetch}~\hs{k}~\hs{=}~\hs{writer}~\hs{(@@getValue}~\hs{store}~\hs{k,}~\hs{[@@k])}
to the task, which in addition to fetching a value from the store, tracks the
corresponding key. Here is an example of \hs{track}ing monadic tasks:

>>>>

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
import Control.Monad.Writer
\end{minted}
\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{haskell}
track :: Task Monad k v -> (k -> v) -> k -> Maybe (v, [k])
track task fetch = fmap runWriter . task (\k -> writer (fetch k, [k]))
\end{minted}
\vspace{1mm}

\noindent
\todo{AM}{Type of track has changed - we no longer supply the Store, just an access fuction}
\todo{AM}{trackM is essential for Shake, what do we do there?}
The implementation uses the standard Haskell \hs{Writer} monad, which allows us
to record addition information -- a list of keys of type \hs{[@@k]} -- when
computing a task. We instantiate \hs{f} to \hs{Writer}~\hs{[@@k]}~\hs{v}, and feed
\hs{fetch}~\hs{k}~\hs{=}~\hs{writer}~\hs{(@@getValue}~\hs{store}~\hs{k,}~\hs{[@@k])}
to the task, which in addition to fetching a value from the store, tracks the
corresponding key. Here is an example of \hs{track}ing monadic tasks:

### 2018-03-16 paper/5-implementations.tex
Redo the Shake code, flip some arguments

@@

The Shake approach for dependency tracking involves recording traces and verifying them. Using the \hs{Trace} type defined in \ref{sec-verifying-traces} we can implement:

\begin{minted}{haskell}
shake :: Hashable v => Build Monad [Trace k v] k v
shake = recursive $ \k _ fetch act -> do
    poss <- traceMatch (\k v -> (==) v . getHash <$> fetch k) k =<< getInfo
    h <- getStoreHash k
    when (h `notElem` poss) $ do
        (dk, v) <- act
        putStore k v
        dh <- mapM getStoreHash dk
        modifyInfo (Trace k (zip dk dh) (getHash v) :)
\end{minted}

@@

\begin{minted}{haskell}
traceMatch :: (Monad m, Eq k) => (k -> Hash v -> m Bool) -> k -> [Trace k v] -> m [Hash v]
traceMatch check k ts = mapMaybeM f ts
    where f (Trace k2 dkv v) = do
                b <- return (k == k2) &&^ allM (uncurry check) dkv
                return $ if b then Just v else Nothing
\end{minted}

@@

\begin{minted}{haskell}
recursive
    :: (m ~ M i k v, Default i)
    => (k -> Maybe [k] -> (k -> m v) -> m ([k], v) -> m ())
    -> Build Monad i k v
recursive step compute = runM . ensure
    where
        ensure k = do
            let fetch x = ensure x >> getStore x
            done <- getTemp
            when (k `Set.notMember` done) $ do
                modifyTemp $ Set.insert k set
                case track compute fetch k of
                    Nothing -> return ()
                    Just act -> step k (getDependenciesMaybe compute k) fetch act
\end{minted}

@@

>>>>

\todo{AM}{Note that anything updating input files must also update Make's info.}
\todo{AM}{Say that in reality that build rules mutate the state themselves.}

@@

The \Shake approach for dependency tracking involves recording traces and verifying them, for which we use the \hs{Trace} type defined in \ref{sec-verifying-traces}. Complete code is given in \S\ref{fig-shake-implementation}.

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Determine whether a trace is relevant to the current state
traceMatch :: (Monad m, Eq k) => (k -> Hash v -> m Bool) -> k -> [Trace k v] -> m [Hash v]
traceMatch check key ts = mapMaybeM f ts
    where f (Trace k dkv v) = do
                b <- return (key == k) &&^ allM (uncurry check) dkv
                return $ if b then Just v else Nothing
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Recursive dependency strategy
recursive
    :: Eq k => (k -> (k -> State (Store i k v, [k]) v) -> State (Store i k v, [k]) (v, [k]) -> State (Store i k v, [k]) ())
    -> Build Monad i k v
recursive step task key store = fst $ execState (ensure key) (store, [])
    where
        ensure key = do
            let fetch k = do ensure k; gets (getValue k . fst)
            s <- get
            when (key `notElem` snd s) $ do
                modify $ \(s, done) -> (s, key:done)
                case trackM task fetch key of
                    Nothing -> return ()
                    Just act -> step key fetch act
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Shake build system
shake :: (Eq k, Hashable v) => Build Monad [Trace k v] k v
shake = recursive $ \key fetch act -> do
    traces <- gets (getInfo . fst)
    poss <- traceMatch (\k v -> (==) v . hash <$> fetch k) key traces
    current <- gets (getHash key . fst)
    when (current `notElem` poss) $ do
        (v, ds) <- act
        modify $ \(s, done) ->
            let t = Trace key [(d, getHash d s) | d <- ds] (getHash key s)
            in (putInfo (t : getInfo s) (putValue key v s), done)

\end{minted}
\vspace{-2mm}
\caption{An implementation of \Shake using our framework.}\label{fig-shake-implementation}
\vspace{-2mm}
\end{figure}

@@

@@

@@

\subsection{Other build systems}

We have implemented the complete product, reusing the \hs{topological} functions etc.
